Journey: A Study of Scalability of Memory Management in Linux

### 2017-05-06 paper.tex
a few minor fixes

When a process maps a region, a \texttt{vm\_area\_struct} is created, but no
memory is allocated. When a process first touches a page, a page fault occurs
and the processor traps to a page fault handler, which allocates memory and
sets up a page table entry for a valid access. This is called \texttt{demand paging}.

@@

On Linux, the system may have multiple swap spaces in the form of swap disk
partitions or swap files on a filesystem. Both have a similar format on disk,
in which the space is broken into a number of swap slots large enough for one
4K page. For each of these swap spaces, the kernel keeps a swap map, which records
the usage of each swap slot so that it can find and allocate them quickly.

@@

The PFRA runs primarily as a kernel thread called kswapd. kswapd runs at
regular intervals and sleeps in between. Also, it is invoked if the system
detects that memory has gone below a threshold such as during a page fault.
kswapd will continue to run until enough pages are freed to bring the free
memory pool above some threshold before going to sleep \cite{utlk}.  This may
cause the faulting process to block for a while. One concern with systems that
have a large memory usage is that kswapd may expend significant CPU time
scanning through hundreds of thousands of pages.

@@

\begin{itemize} \item \textit{Control}. This workload is only run for the memory
overhead experiments. It simply prints the $\Delta_i$ for each $i$ while performing nops. It's graph
is a straight line with positive slope. This memory usage results from the
buffer containing stdout, where our readings are being printed. All of our
experiments have the same stdout overhead, so this benchmark serves a baseline
for comparison.

@@

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/mmap_mem_usage}
    \caption{$\Delta_i$ as discussed in section \ref{ss_vm_area_struct}. Note that
    graphs of \texttt{frag} and \texttt{stride} overlap. Also, \texttt{cont} and
    \texttt{control} overlap.}
    \label{fig:mmap_mem_usage}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \caption {Tail latency of memory accesses}
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
    \label{fig:latency_table}
\end{figure}

>>>>

When a process maps a region, a \texttt{vm\_area\_struct} is created, but no
memory is allocated. When a process first touches a page, a page fault occurs
and the processor traps to a page fault handler, which allocates memory and
sets up a page table entry for a valid access. This is called \textit{demand paging}.

@@

On Linux, the system may have multiple swap spaces in the form of swap disk
partitions or swap files on a filesystem. Both have a similar format on disk,
in which the space is broken into a number of swap slots large enough for one
4KB page. For each of these swap spaces, the kernel keeps a swap map, which records
the usage of each swap slot so that it can find and allocate them quickly.

@@

The PFRA runs primarily as a kernel thread called kswapd. kswapd runs at
regular intervals and sleeps in between. Also, it is invoked if the system
detects that memory has gone below a threshold, such as during a page fault.
kswapd will continue to run until enough pages are freed to bring the free
memory pool above some threshold before going to sleep \cite{utlk}.  This may
cause the faulting process to block for a while. One concern with systems that
have a large memory usage is that kswapd may expend significant CPU time
scanning through hundreds of thousands of pages.

@@

\begin{itemize} \item \textit{Control}. This workload is only run for the memory
overhead experiments. It simply prints the $\Delta_i$ for each $i$ while performing nops. Its graph
is a straight line with positive slope. This memory usage results from the
buffer containing stdout, where our readings are being printed. All of our
experiments have the same stdout overhead, so this benchmark serves a baseline
for comparison.

@@

\begin{figure}[t]
    \includegraphics[width=\columnwidth]{figures/mmap_mem_usage}
    \caption{$\Delta_i$ as defined in section \ref{ss_vm_area_struct}. Note that
    graphs of \texttt{frag} and \texttt{stride} overlap. Also, \texttt{cont} and
    \texttt{control} overlap.}
    \label{fig:mmap_mem_usage}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
    \caption {Tail latency of memory accesses.\label{fig:latency_table}}
\end{figure}
